### 2024-10-07 Manuscript/main_ecography_R1.tex
Update main_ecography_R1.tex

For the decreasing size, increasing density and shifting space stressors, we estimated the parameters $\rho$, $D$ and $\rho'$ recursively with the \texttt{dispRity} package to obtain the required amount of data to be removed (\texttt{dispRity::reduce.space}, \citealt{guillerme2018disprity,guillerme2020shifting}).
Note that the algorithm used here to simulate the stressors depend, to some extent, to the distribution of the data for each trait space.
The algorithms used can lead to similar stressors effects depending on the simulated data distribution.
For example, ``increasing density'' and ``decreasing size'' stressors will lead to just what their name suggest if the observations are uniformly distributed (just average increase in density for ``increasing density'' and just overall decrease in size for the ``decreasing size'' stressors) but will have more or less similar effects (albeit no identical) in the case of a normal distribution (average increase in density and overall decrease in size for both ``increasing density'' and ``decreasing size'' stressors.)
These steps resulted in 3420 simulated trait spaces (171 simulated spaces $\times$ 5 stressors $\times$ 4 intensities of data removal).
We used 171 replicates per simulation because that was empirically the smallest number of simulations required to reach a variance between simulations lower than 1\% across all metrics (i.e. any additional simulation added less that 1\% extra variance).
See Table \ref{Tab:mechanisms} for a biological description of these stressors and Figure \ref{Fig:simulations} for a visual description of them.

@@

We repeated each simulation pipeline (generate a trait space and apply the stressor) for uncorrelated 4 dimensional traits (Figure \ref{Fig:simulation_results}) and uncorrelated 2 and 8 dimensional traits (supplementary materials).
We limited our simulations to a relatively small number of dimensions due to the constraints of some of the metrics used (e.g., \texttt{TPD::TPDsMean} is only implemented for up to 4 dimensions; \citealt{carmona2019trait}) but also to avoid dealing with the curse of dimensionality \citep{bellman1957dynamic}.
This curse changes the properties of space occupancy in a non-linear way depending on the observation's distribution and the number of dimensions.
For example, the volume of a trait space (or hypervolume when using more than 3 dimensions) typically tends to zero in a high number of dimensions.
However, the rate at which it approaches zero is not linear and depends on the distribution of each observation on every dimension.
This makes it practically difficult to compare two randomly generated spaces with similar characteristics (e.g. for two spaces with 200 observations and 10 dimensions generated in the exact same way, one might have a hypervolume of nearly 0 and the other one of $10^5$).
Note also that in our simulations, all dimensions have the same properties (i.e. same variance and distribution).
This is often not the case in empirical cases (e.g. see our empirical example) where the dimensions have a decreased variance due to ordination techniques.

\begin{table}
\center
\scriptsize
\begin{tabular}{p{0.3\linewidth}|p{0.3\linewidth}|p{0.4\linewidth}}
\textbf{Approximated mechanism (biological)} & \textbf{Used mechanism (algorithm in \texttt{dispRity::reduce.space})} & \textbf{Stressor description}\\
\hline
  Null mechanism & Random removal (``random'') & No overall change in species traits or community structure except for the reduction in the number of observations.\\
  Equalising fitness & Size change (``size'') & When species with trait combinations that are more extreme (i.e. species located away from the centre of the trait space) are disadvantaged due to some resource concentration gradient. \\
  Facilitation & Density change (``density'') & When species with closely shared trait combinations (i.e. located close together in the trait space) are more likely to survive the stressor. In other words, this simulates the idea that some species are more likely to be resilient if they are present in a community of species with shared traits rather than the opposite.\\
  Environmental filtering & Position change (``position'') & When an increase in species trait similarity happens through some strong abiotic constraints \citep{cornwell2006trait}. For example, some trait combinations become more and more unlikely due to some environmental constraints (i.e. some regions of the trait space become unsuitable).\\
  Competitive exclusion & Evenness change (``evenness'') & When functionally similar species compete more intensively with one another than with functionally dissimilar species (``competitive exclusion principle''; \citealt{hardin1960competitive}). At the extreme, it implies that only dissimilar individuals will coexist (``limiting similarity principle''; \citealt{macarthur1967limiting})\\

@@

To ease interpretation of the results, we centred the results for the 12 metrics for the 4 non null stressors by the results for the null stressor and then scaled them.
Effectively, for each metric and each single simulation, we subtracted the results under the null stressor to the ones under the other stressors and divided the resulting centred metrics by the maximum value for each metric between each simulation.
This resulted in each metric being scaled between –1 and 1 where 0 corresponds to no change compared to the null and 1 and –1 the maximum positive and negative deviations relative to the null.
In parallel, we also compared the distribution of the 171 scores for each metric with each reduction level for each non-null stressor to the null stressor (with no scaling or centering) using pairwise t-tests.
The results of these tests are displayed in figure \ref{Fig:simulation_results} (for the non-null stressor, grey distributions are not clearly distinguishable from the null stressor).

>>>>

For the decreasing size, increasing density and shifting space stressors, we estimated the parameters $\rho$, $D$ and $\rho'$ recursively with the \texttt{dispRity} package to obtain the required amount of data to be removed (\texttt{dispRity::reduce.space}, \citealt{guillerme2018disprity,guillerme2020shifting}).
Note that the algorithm used here to simulate the stressors depend, to some extent, to the distribution of the data for each trait space.
The algorithms used can lead to similar stressors effects depending on the simulated data distribution.
For example, ``increasing density'' and ``decreasing size'' stressors will lead to just what their name suggest if the observations are uniformly distributed (just average increase in density for ``increasing density'' and just overall decrease in size for the ``decreasing size'' stressors) but will have more or less similar effects (albeit no identical) in the case of a normal distribution (average increase in density and overall decrease in size for both ``increasing density'' and ``decreasing size'' stressors.)
These steps resulted in 3420 simulated trait spaces (171 simulated spaces $\times$ 5 stressors $\times$ 4 intensities of data removal).
We used 171 replicates because that was empirically the smallest number of replicates required to reach a variance between replicates lower than 1\% across all metrics (i.e. any additional replicates beyond 171 added less that 1\% extra variance).
See Table \ref{Tab:mechanisms} for a biological description of these stressors and Figure \ref{Fig:simulations} for a visual description of them.

@@

We repeated each simulation pipeline for 4 dimensional traits (generate a trait space and apply the stressor - Figure \ref{Fig:simulation_results}) for 2 and 8 dimensional traits with not correlation between each dimensions in each simulation pipeline (supplementary materials).
We limited our simulations to a relatively small number of dimensions due to the constraints of some of the metrics used (e.g., \texttt{TPD::TPDsMean} is only implemented for up to 4 dimensions; \citealt{carmona2019trait}) but also to avoid dealing with the curse of dimensionality \citep{bellman1957dynamic}.
This curse changes the properties of space occupancy in a non-linear way depending on the observation's distribution and the number of dimensions.
For example, the volume of a trait space (or hypervolume when using more than 3 dimensions) typically tends to zero in a high number of dimensions.
However, the rate at which it approaches zero is not linear and depends on the distribution of each observation on every dimension.
This makes it practically difficult to compare two randomly generated spaces with similar characteristics (e.g. for two spaces with $200$ observations and $10$ dimensions generated in the exact same way, one might have a hypervolume of nearly $0$ and the other one of $10^5$).
Note also that in our simulations, all dimensions have the same properties (i.e. same variance and distribution).
This is often not the case in empirical cases (e.g. see our empirical example) where the dimensions have a decreased variance due to ordination techniques.

\begin{table}
\center
\scriptsize
\begin{tabular}{p{0.3\linewidth}|p{0.3\linewidth}|p{0.4\linewidth}}
\textbf{Approximated mechanism (biological)} & \textbf{Used mechanism (algorithm in \texttt{dispRity::reduce.space})} & \textbf{Stressor description}\\
\hline
  Null mechanism & Random removal (``random'') & No overall change in observations' traits or community structure except for the reduction in the number of observations.\\
  Equalising fitness & Size change (``size'') & When observations with trait combinations that are more extreme (i.e. observations located away from the centre of the trait space) are disadvantaged due to some resource concentration gradient. \\
  Facilitation & Density change (``density'') & When observations with closely shared trait combinations (i.e. located close together in the trait space) are more likely to survive the stressor. In other words, this simulates the idea that some observations are more likely to be resilient if they are present in a community of observations (e.g. a community of species) with shared traits rather than the opposite.\\
  Environmental filtering & Position change (``position'') & When an increase in observations' trait similarity happens through some strong abiotic constraints \citep{cornwell2006trait}. For example, some trait combinations become more and more unlikely due to some environmental constraints (i.e. some regions of the trait space become unsuitable).\\
  Competitive exclusion & Evenness change (``evenness'') & When functionally similar observations compete more intensively with one another than with functionally dissimilar observations (``competitive exclusion principle''; \citealt{hardin1960competitive}). At the extreme, it implies that only dissimilar individuals will coexist (``limiting similarity principle''; \citealt{macarthur1967limiting})\\

@@

To ease interpretation of the results, we centred the results for the 12 metrics for the 4 non null stressors by the results for the null stressor and then scaled them.
Effectively, for each metric and each single replicate, we subtracted the results under the null stressor to the ones under the other stressors and divided the resulting centred metrics by the maximum value for each metric between each replicate.
This resulted in each metric being scaled between –1 and 1 where 0 corresponds to no change compared to the null and 1 and –1 the maximum positive and negative deviations relative to the null.
In parallel, we also compared the distribution of the 171 scores for each metric with each reduction level for each non-null stressor to the null stressor (with no scaling or centering) using pairwise t-tests.
The results of these tests are displayed in figure \ref{Fig:simulation_results} (for the non-null stressor, grey distributions are not clearly distinguishable from the null stressor).
